Response diversity is a major driver of temporal stability in complex food webs

### 2024-08-29 paper/manuscript.Rmd
Cleaning targets pipeline (removing old code not used)

---
title: Response diversity and food-web structure drive stability and stability-richness relationships in complex communities
author: "Alain Danet (1), Sonia Kéfi (2,3), Thomas F Johnson (1), Andrew P. Beckerman (1)"
date: ""
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
  officedown::rdocx_document:
    number_sections: false
    base_format: "bookdown::word_document2"
always_allow_html: true
bibliography: references.bib
link-citations: true
csl: nature.csl
#csl: ecology-letters.csl
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
---

@@

# Abstract (179/200 words)

Global environmental change constitutes a suite of major threats to biodiversity
and ecosystem functioning. These threats can materialised via changes to the
stability of communities and the services they provide. Despite independent
advances in understanding the impact on stability of community size and
structure, synchrony and responses to environmental variation, there remains no
core theory connecting these key processes. Developing this theory requires
combining concepts from diversity - stability relationships, synchrony and the
portfolio effect, and the emerging field of response diversity. Using a
stochastic Bioenergenetic food-web model of 20-50 interacting species, we show
that response diversity and species richness are also strong drivers of overall
community stability in complex food-webs. We further show that connectance and
average trophic level consistently dampened asynchrony but that average trophic
level stabilises population stability. Positive stability-richness relationships
always emerged only in presence of response diversity, but interaction strength,
connectance, average trophic level strongly interact with response diversity to
determine the sign of stability-diversity relationships. Our study disentangles
the complex effects of food-web structure on community stability and its
pathway, showing that response diversity is a mechanism that generates positive
stability-diversity relationships, beyond plant communities.

@@

Ecological communities are facing ever-growing environmental variability caused
by climate and environmental change[@lee_ipcc_2023;
@simmons_refocusing_2021]. Decreases in the temporal stability of communities
are increasingly documented[@olivier_urbanization_2020;
@de_boeck_patterns_2018; @hansen_climate_2013; @bluthgen_land_2016; @harris_biological_2018], raising
concerns about the long-term consequences at the ecosystem scale. Repeated calls
have been made to clarify the role of environmental stochasticity in driving the
temporal stability of ecological communities[@yang_predictability_2019]. It
remains challenging because species have a diversity of response to
environmental change [@elmqvist_response_2003; @loreau_species_2008;
@ives_stability_2000; @ross_how_2023] (“response diversity”) and they are
entangled in complex communities such as food-webs, that both ultimately drive
community stability[@may_will_1972; @angelis_stability_1975; @mccann_weak_1998;
@mccann_reevaluating_1997; @thebault_stability_2010] and stability-richness
relationships[@thebault_trophic_2005; @thebault_relationship_2006]. Despite the
overarching effects of response diversity and food-web structure on temporal
community stability, they have been so far studied in isolation.

The temporal stability of ecological communities stems from average population
stability and asynchrony of population fluctuations[@thibaut_understanding_2013;
@zhao_biodiversity_2022]. While higher environmental stochasticity/variability
decrease population stability by increasing the amplitude species oscillations,
higher species response diversity increases asynchrony of species
fluctuations[@doak_statistical_1998; @loreau_species_2008;
@thibaut_understanding_2013]. This is increasingly well-established in
simplified species assemblages, where high response diversity increases
asynchrony and is the determinant of positive stability-diversity relationships
across many taxa[@tilman_biodiversity_2006; @olivier_urbanization_2020;
@zhao_biodiversity_2022]. However, it remains largely unknown how the effects of
environmental stochasticity and response diversity on population stability and
asynchrony are mediated in the context of complex communities, such as
food-webs.

Complexity-stability theory well documented that the structure of food-webs
determines population stability [@may_will_1972; @angelis_stability_1975;
@mccann_weak_1998; @mccann_reevaluating_1997; @vasseur_environmental_2007;
@eschenbrenner_diversity_2023], and stability-richness relationships
[@thebault_trophic_2005; @thebault_relationship_2006]. Population stability is
indeed documented to decrease with increasing interaction strength
[@may_will_1972; @mccann_weak_1998], connectance [@may_will_1972], but to be
higher at higher trophic level [@shanafelt_stability_2018; @danet_species_2021].
In result, the distribution and strength of species interactions, as well as
trophic level might have a strong effect on community stability and
stability-richness relationships [@thebault_trophic_2005;
@thebault_relationship_2006]. This expectation is reinforced by documented
examples where predators can mediate the effect of response diversity on
asynchrony [@vasseur_environmental_2007], and synchronise the dynamics of their
prey [@raimondo_interspecific_2004; @korpimaki_predatorinduced_2005]. Recent
study in plant communities suggests that asynchrony stems mainly from the
statistical averaging of independent species fluctuations (“portfolio effects”),
rather than compensatory dynamics resulting from response diversity and species
interactions [@zhao_biodiversity_2022]. But we might expect that prey switching
and trophic cascades arising in food-webs lead to higher compensatory dynamics
than in plant communities [@mccann_diversitystability_2000;
@fahimipour_compensation_2017].

Here, we investigated the effects of response diversity and food-web structure
on community stability and on stability-richness relationships in stochastic
environment. To do so, we developed an extension of the Bioenergetic food web
model [@brose_allometric_2006; @lajaaiti_ecologicalnetworksdynamicsjl_2024]
which includes environmental stochasticity and response diversity
[@loreau_species_2008; @vasseur_environmental_2007; @ripa_food_2003]. We
included environmental stochasticity on allometric natural death rate, such as
species of higher trophic level had lower death rate in
average[@vasseur_environmental_2007], response diversity was included as the
inverse of species correlation in their mortality rates.  We first decipher the
respective effects of food-web structure, environmental stochasticity, and
response diversity on community stability and its partition into population
stability and asynchrony (i.e. portfolio and compensatory effects). We then
investigate how food-web structure and response diversity interact to determine
the sign of stability-richness relationship.

@@

```{r}
sem_cap <- paste0(
  "Structural Equation Model linking food-web structure and stressors to temporal
  stability of community biomass. a, continuous red arrows display positive and
  dashed blue arrows display negative standardised effects, width of the arrows
  being proportional to the absolute values of the standardised effects. Only
  the standardised coefficients whose absolute values are superior to 0.05 are
  displayed. The full table of coefficients including control variables is
  displayed in Table S4. A more complete SEM containing further partition of
  compensatory effects and control variables is displayed in Table S5. b, total
  effects of food-web structure on community stability, asynchrony, and
  population stability. N = ", nrow(tar_read(sim_fw)), " simulations.") %>%
  str_replace_all("\\s+", " ")
```

```{r sem, fig.cap=sem_cap, fig.id = "sem"}
knitr::include_graphics(here("report", "figures", "p_sem_fig.pdf"))
```

@@

## Drivers of community stability, population stability and asynchrony

@@

rac <- get_eff(y = "async", x = "cpe")
ras <- get_eff(y = "async", x = "sae_total")
```

The `r format(m_sim(col = "n"), scientific = FALSE)` simulations generated a
wide range of food-web structure (Median (5%, 95%), species richness:
`r m_sim("richness")`, connectance: `r m_sim("ct_alive")`, average interaction
strength: `r m_sim("avg_int_strength")`, maximum trophic level:
`r m_sim("max_tlvl")`). We quantified the  (1) the respective effects of
population
stability and asynchrony on community stability, (2) the source of asynchrony,
either portfolio effects or compensatory dynamics, and disentangled (3) how
stressors (environmental stochasticity, response diversity) and food-web
structure (species richness, average interaction strength, average trophic
level, connectance) determine average population stability, portfolio effects
(using Doak's definition), and compensatory dynamics stemming from species
interactions. We did so by using a structural equation model which included all
the simulation variables as control (i.e. predator interference, predator-prey
mass ratio) to avoid the detection of spurious non-causal effects (See Methods,
Table/Figure SXX). Overall, we found that population stability had a positive
effect twice higher than asynchrony on the temporal stability of community
biomass (resp. $r_\delta$= `r rscp` and `r rsca`, $r_\delta$ being the
standardised coefficient, Fig. \@ref(fig:sem)a). We further found that
asynchrony was driven more by portfolio effects than by compensatory dynamics
stemming from species interactions (resp. $r_\delta$= `r rac` and `r ras`).
However, we found opposite results, i.e. that compensatory dynamics contributed
more than portfolio effects to asynchrony, using an alternative portfolio
effects definition that attributes the effects of response diversity to
compensatory effects instead of portfolio effects (Fig. SXX).

Food-web structure had consistent negative effects on asynchrony, mainly through
portfolio and compensatory effects (Fig. \@ref(fig:sem)a). Indeed, average
trophic level and connectance had negative effects on the compensatory effects
stemming from species interactions (resp. $r_\delta$= `r rct` and `r rcc`, Fig.
\@ref(fig:sem)a), while average interaction strength had a small negative
effects on portfolio effects ($r_\delta$= `r rsi`). In contrast, response
diversity and species richness had strong positive effects on asynchrony,
mainly through portfolio effects. Response diversity had a strong positive
effect on portfolio effects ($r_\delta$= `r rsd`), and a negative one on
compensatory dynamics ($r_\delta$= `r rcd`). In turn, species richness had a
positive effect on both portfolio and compensatory effects (resp.
$r_\delta$=`r rsr` and `r rcr`). Finally, environmental stochasticity had a
small negative effect on compensatory dynamics.

In turn, we found little effect of food-web structure on population stability,
except for average trophic level having a strong positive effect on
population stability ($r_\delta$= `r rpt`, Fig. \@ref(fig:sem)a). In turn, the
variance of environmental stochasticity had a strong negative effect on
population stability ($r_\delta$= `r rpe`), while response diversity had a small
positive effect on population stability ($r_\delta$= `r rpd`).

@@

Deriving the sum of direct and indirect effects from the structural equation
model, we show that both the variance of environmental stochasticity and species response
diversity had by far the most important effects on community stability (resp.
$r_\delta$= `r rtsenv` and `r rtsdiv`, Fig. \@ref(fig:sem)b). Species richness
and average trophic level had similar positive total effects on community stability
($r_\delta$=`r rtsr`), but through opposite pathways and antagonistic effects on
asynchrony and population stability. While species richness increased community
stability through a strong positive effect on asynchrony ($r_\delta$=`r rtar`), it
also weakly decreased population stability ($r_\delta$=`r rtpr`). At the opposite, average
trophic level increased community stability through strong positive effect on
population stability ($r_\delta$=`r rtpt`), but dampened by a negative total
effect on asynchrony ($r_\delta$=`r rtat`). Interestingly, average interaction
strength also had a positive effect on population stability and a negative one
on asynchrony (resp. $r_\delta$= `r rtpi` and `r rtai`), but resulting in a
total negative effect on community stability ($r_\delta$= `r rtsi`).
Connectance, instead, had a negative effect on community stability ($r_\delta$=
`r rtsct`) through consistent negative effects on asynchrony and population
stability (resp. $r_\delta$=`r rtact` and `r rtpct`).

## Stability-richness relationship

@@

)
pred_cap <- paste0("Slope of the community stability - species
  richness relationship according to food-web structure and response diversity. (a),
  Diagrams displaying the slope coefficients of the effect of species richness
  on community stability. The average values of the control variables were used
  to generate the predictions: ", pv_cap, ". (b, c): Examples of predictions
  from the model for two combinations of average trophic level and response
  diversity (values displayed in panel (a)). Lines display the mean predictions
  of the model. The coefficients of the linear model are displayed in Fig. S4.
  Relationships between population stability, asynchrony and species richness
  are displayed in Fig. S5."
  ) %>%
  str_replace_all("\\s+", " ")
```

@@

Using a more complex linear model including interactions among food-web
structure, response diversity and environmental stochasticity, we found that
food-web structure strongly interacts with response diversity to determine the
sign of the stability-richness relationships. In the absence of response
diversity, we found only negative stability-richness, regardless of the food-web
structure (Fig. \@ref(fig:pred)a). Response diversity enabled the rise of
positive stability-richness relationships, enhanced by average interaction
strength and dampened by average trophic level and connectance. Higher
interaction strength led to positive stability-richness relationships for lower
values of response diversity, interaction strength enhanced asynchrony-richness
relationships (Fig. SXX), so that the stronger positive stability-richness
relationships were found at both high interaction strength and response
diversity. In contrast, higher average trophic level and connectance both led to
negative stability-richness relationships, because they led to more negative
population stability-richness relationships and weaker positive
asynchrony-richness relationships (Fig. SXX). Overall, our results highlight
strong contrast between the relatively small effects of food-web structure on overall
community stability (Fig. \@ref(fig:sem)b) and their strong effects on
stability-richness relationships (Fig. \@ref(fig:pred)a).

@@

Temporal community stability is well documented in simplified assemblages
such as plant communities, where community stability is driven by asynchrony in
species fluctuations. In contrast, our results show that community stability in
complex food-webs was mainly driven population stability rather than asynchrony.
In the range of our simulations, we identified that environmental variability
and response diversity were the main driver of community stability, far more
than any food-web structure component, a result consistent across methodological
choices (Fig. SXX). Moreover, we found that species richness and food-web
structure have antagonistic but consistent effect on community stability.
Finally, we found that the sign of the stability-richness relationship is
strongly mediated by the interaction between food-web structure and response
diversity. These results highlight the need to account for both food-web
structure and response diversity to improve our understanding of the stability
of ecological communities in variable environments.

Our results show that community stability in complex food-webs is mainly driven
by population stability rather than synchrony, which contrasts with empirical
data that reported an equivalent influence of population stability and
asynchrony on community stability in birds and bats assemblages
[@olivier_urbanization_2020]. However, our results are in accordance with
previous findings in empirical food-webs [@danet_species_2021] and
food-web models [@eschenbrenner_diversity_2023]. Are there fundamental
differences between single trophic and food-web in the contribution of
population stability and asynchrony? We found that all food-web metrics, i.e.
average trophic level, connectance, and average interaction strength all
decreased asynchrony. In line with previous studies showing that predators can
synchronise their preys [@raimondo_interspecific_2004;
@kortsch_disentangling_2021], and that predator-prey couples are more
synchronous [@danet_species_2021], those results suggest that trophic links
might dampen overall asynchrony in ecological communities. A complementary
explanation could be related to the biomass structure of food-webs. Food-webs
typically includes range of bodymasses much larger than in species assemblages,
differing in bodymass by several orders of magnitude [@brose_predator_2019].
This highly uneven biomass distribution can translate in uneven species
variability, thereby creating uneven contributions of species to asynchrony
[@thibaut_understanding_2013] and limits portfolio effects and compensatory
dynamics in species rich communities [@zhao_biodiversity_2022]. The fact that
average trophic level, connectance, and average interaction strength all
decreased evenness of portfolio effects lends toward this hypothesis. Our
results then reinforce the evidence that community stability in food-web might
be driven more by population stability than asynchrony.

We further found that asynchrony was driven more by portfolio effects than
compensatory dynamics. But this results was dependant of the definition of
portfolio effects, as we found the opposite while taking Tilman's
definition[@tilman_ecological_1999; @zhao_biodiversity_2022] instead of Doak's
one[@doak_statistical_1998], in contrast with previous assessment in
experimental plant communities[@zhao_biodiversity_2022]. Portfolio effects is
supposed to measure stabilising effects of species richness in the absence of
species interactions[@loreau_biodiversity_2021]. While Tilman's considers that
portfolio effects are emerging from statistical averaging of independent species
fluctuations alone (i.e. in the absence of covariance among species), Doak's
explicitly states that statistical averaging can only emerge if compensatory
dynamics hold by response diversity is present[@zhao_biodiversity_2022]. While
the definition of portfolio effects is a long-standing
debate[@zhao_biodiversity_2022], our results show that the importance of
portfolio effects depends on where response diversity acts. In agreement with
previous studies[@loreau_species_2008; @doak_statistical_1998;
@thibaut_understanding_2013; @loreau_biodiversity_2021], we believe that the
definition of portfolio effects standing from statistical averaging and
environmental compensatory dynamics generated by response diversity is the more
appropriate as the absence of species interactions does not guarantee independent
species fluctuations, because common environment tend to generate positive
covariances among species[@loreau_biodiversity_2021]. The fact that
environmental stochasticity alone had a really small effect on asynchrony to the
difference of response lends toward this interpretation. So, our study suggests
that response diversity is a key ingredient to generate asynchrony, not only in
plant communities, but also in food-webs. We further show that explicitly
manipulating response diversity highlight that it is the main driver of
asynchrony, and that the respective contribution of portfolio effects and
compensatory dynamics to asynchrony depends on the decision to consider that
response diversity is part of portfolio effects or compensatory dynamics.

>>>>

---
title: Response diversity is a major driver of temporal stability in complex food webs
author: "Alain Danet (1), Sonia Kéfi (2,3), Thomas F. Johnson (1), Andrew P. Beckerman (1)"
date: ""
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
  officedown::rdocx_document:
    number_sections: false
    base_format: "bookdown::word_document2"
always_allow_html: true
bibliography: references.bib
link-citations: true
csl: nature.csl
#csl: ecology-letters.csl
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
---

@@

# Abstract

Global change constitutes a suite of major threats to biodiversity and ecosystem
functioning. These threats can materialise via changes in the temporal stability
of ecological communities and the services they provide. However, the majority
of research on stability has focused on single trophic level communities and has
not yet integrated classic theory about species richness and food web structure
with more recent theory centred on response diversity and stochasticity. Using a
stochastic bioenergenetic food-web model, we reveal that response diversity is a
major driver of community stability. Moreover, positive stability-richness
relationships emerge only in the presence of response diversity. In contrast to
previous work,  food-web structural properties are only secondary drivers of
overall community stability, but interact with response diversity to determine
the sign of the stability-richness relationship. Our study highlights the
complex pathways by which food-web structure and response diversity drive
community stability, and raises concerns about how the loss of response
diversity may lead to a breakdown of stability and the capacity for these
communities to deliver functions and services to human societies.

@@

Global change, characterised by multiple, simultaneous threats to
biodiversity[@lee_ipcc_2023; @simmons_refocusing_2021], is having impacts on the
temporal stability of complex ecological communities[@olivier_urbanization_2020;
@de_boeck_patterns_2018; @hansen_climate_2013; @bluthgen_land_2016;
@harris_biological_2018]. A breakdown of stability impacts the capacity for
these communities to deliver functions and services to human
societies[@loreau_ecological_2022]. Predicting and managing the stability and
services provided by biodiverse communities requires reconciling multiple
concepts and theories about temporal stability in species diverse and
structurally complex communities. These include classic ideas about the role of
trophic structure and interaction strengths[@may_will_1972;
@angelis_stability_1975; @mccann_weak_1998; @mccann_reevaluating_1997;
@thebault_trophic_2005; @thebault_relationship_2006; @thebault_stability_2010],
newer concepts of response diversity, average population stability, asynchrony
and portfolio effects and integrative theories about the relationship between
stability, structure and diversity.

An abundant literature has established that species diversity can buffer the
effect of environmental perturbations on ecological communities, thereby
increasing the temporal stability (hereafter “stability”) of
communities[@yachi_biodiversity_1999; @doak_statistical_1998;
@tilman_ecological_1999; @de_mazancourt_predicting_2013;
@thibaut_diversity_2012; @zhao_biodiversity_2022]. Because different species
have diverse niches and environmental preferences, they may react differently to
the same environmental change [@elmqvist_response_2003; @ross_how_2023]. This
diversity of species responses to environmental change, coined as “response
diversity”, has been identified as a crucial mechanism that buffers the response
of ecological communities to environmental perturbations
[@yachi_biodiversity_1999; @thibaut_diversity_2012;
@thibaut_understanding_2013]. For example, in a given community, some species
will respond positively and others negatively to changes in temperature,
depending on their thermal niche [@frund_response_2013]. Furthermore, it has
also been shown recently that the stability of communities can be decomposed
into the average population stability and the asynchrony of population
fluctuations [@thibaut_understanding_2013; @zhao_biodiversity_2022]. Average
population stability quantifies the average variability in species abundances,
while asynchrony refers to how much the temporal fluctuations of the different
species of the community are buffering each other. Recent studies have shown
that asynchrony can be further resolved into “portfolio effects”, the
statistical averaging of independent species fluctuations in response to
environmental perturbations, and “compensatory dynamics”, where one species’
decrease can be compensated for by the increase in another
species[@doak_statistical_1998; @thibaut_understanding_2013;
@zhao_biodiversity_2022].

These mechanisms have been well-established in data from single trophic level
(e.g. competitive) communities, such as grasslands, butterflies, birds, and bats
communities[@de_mazancourt_predicting_2013; @olivier_urbanization_2020]. In
these studies, response diversity increases asynchrony, and is a more important
driver of community stability than population
stability[@olivier_urbanization_2020]. In these same studies, the portfolio
effect was found to play a key role in the emergence of positive
diversity-stability relationships in experimental
grasslands[@zhao_biodiversity_2022].

Despite these data, there is limited evidence and understanding for how these
collections of theories integrate to define stability in food webs with
substantial diversity and vertical complexity (i.e. trophic levels)
[@yang_predictability_2019; @danet_species_2021; @eschenbrenner_diversity_2023;
@srednick_understanding_2024]. Classic theory and concepts about
stability-complexity [@may_will_1972; @angelis_stability_1975],
stability-structure [@mccann_weak_1998; @mccann_reevaluating_1997;
@thebault_trophic_2005; @eschenbrenner_diversity_2023] and stability-richness [@ives_stability_2000;
@thebault_trophic_2005; @thebault_relationship_2006] offer some expectation.
They suggest that population stability is expected to decrease with increasing
interaction strength[@may_will_1972; @mccann_reevaluating_1997;
@mccann_weak_1998] and food web connectance[@thebault_trophic_2005] but to
increase with average trophic level [@shanafelt_stability_2018;
@danet_species_2021]. As a result, the distribution and strength of species
interactions, as well as the average trophic level, can have strong effects on
community stability as well as on stability-richness
relationships[@thebault_trophic_2005; @thebault_relationship_2006]. These
relationships indicate that the structure of food-webs strongly influences
population stability and stability-richness relationships.

The more recent theory and concepts centred on response diversity and asynchrony
add further predictions[@thibaut_understanding_2013; @zhao_biodiversity_2022].
With vertical structure and trophic interactions, consumers can synchronise the
dynamics of their prey[@raimondo_interspecific_2004;
@korpimaki_predatorinduced_2005], can offset the effects of response diversity
on asynchrony and can generate substantial compensatory dynamics via prey
switching and trophic cascades[@mccann_weak_1998; @mccann_diversitystability_2000;
@fahimipour_compensation_2017].

Despite this collection of historical and more modern theory, we don’t know what
determines stability in vertically structured, species rich communities.
Against this gap in understanding, we specifically investigate the effects of
response diversity and food-web structure on community stability and on
stability-richness relationships in a variable  environment. The key question is
whether the new concepts of environmental stochasticity, response diversity and
asynchrony are the main mechanisms driving community stability versus more
classic concepts tied to food-web structure and diversity.

We explore this key question by extending  the  bioenergetic food web
model[@brose_allometric_2006; @yodzis_body_1992], to incorporate environmental
stochasticity and response diversity[@vasseur_environmental_2007;
@ives_stability_2007; @ripa_food_2003]. We introduced environmental
stochasticity by making species death rates
stochastic[@vasseur_environmental_2007]. Response diversity is
introduced as the correlation of stochasticity among species, so that species
that experience environmental changes in the same way show positively correlated
variations in their death rates.

We generated over 45000 in-silico communities with diverse richness, trophic
levels and connectance and experimentally simulated variable degrees of
interaction strength, response diversity and environmental stochasticity.  For
each community, we evaluated the effects on stability of species richness,
response diversity, community structure and stochasticity.  Stability was itself
partitioned in our analyses into population stability and asynchrony (i.e.
portfolio and compensatory effects). Finally, we also evaluated the occurrence
of negative and positive relationships between community stability and species
richness.

@@

```{r}

sem_cap <- paste0(
  "Structural Equation Model linking food-web structure and stressors to the
  temporal stability of community biomass. a, Continuous blue arrows display
  positive and dashed red arrows display negative standardised effects, width of
  the arrows being proportional to the absolute values of the standardised
  effects. Only the standardised coefficients whose absolute values are superior
  to 0.05 are displayed. The full table of coefficients including control
  variables is displayed in Table S2. b, Total effects of food-web structure on
  community stability, asynchrony, and population stability derived from the
  structural equation model. N = ", nrow(tar_read(sim_fw)), " simulations.") %>%
  str_replace_all("\\s+", " ")
```

```{r sem, fig.cap=sem_cap, fig.id = "sem"}
knitr::include_graphics(here::here("report", "figures", "p_sem_fig.pdf"))
```

@@

@@

rac <- get_eff(y = "async", x = "cpe_int")
ras <- get_eff(y = "async", x = "sae_doak")
```

Applying a structural equation model on the outcomes of food web simulations
(see Methods) revealed that the temporal stability of community biomass is
positively influenced by both population stability and asynchrony, and that
population stability has twice the effect of asynchrony (resp. $r_\delta$=
`r rscp` and `r rsca`, $r_\delta$ being the standardised coefficient, Fig.
\@ref(fig:sem)a).

We further found that asynchrony is driven more by portfolio effects than by
compensatory dynamics (resp. $r_\delta$= `r ras` and `r rac`). In more details, response
diversity has strong positive effects on asynchrony, mainly through portfolio
effects ($r_\delta$= `r rsd`), but also through a weaker positive effect on compensatory
dynamics ($r_\delta$= `r rcd`). All other studied properties were found to have relatively
much weaker effects on asynchrony. Food web structural properties – namely
average trophic level, connectance and average interaction strength – have
consistent negative effects on asynchrony, through both portfolio and
compensatory effects (Fig. \@ref(fig:sem)a). Indeed, as average trophic level and connectance
increase compensatory effects are reduced (resp. $r_\delta$= `r rct` and `r rcc`, Fig.
\@ref(fig:sem)a), while average interaction strength has a small negative effect
on portfolio effects ($r_\delta$= `r rsi`). Species richness has a positive
effect on both portfolio and compensatory effects (resp. $r_\delta$=`r rsr` and
`r rcr`). Finally, as expected, environmental stochasticity has little effect on
asynchrony.

Focusing on population stability, environmental stochasticity was found to have
a strong negative effect on it ($r_\delta$= `r rpe`). All other studied properties were
found to have little effect on population stability, except for the average
trophic level which has a small positive effect ($r_\delta$= `r rpt`, Fig.
\@ref(fig:sem)a, but see Fig. S1) and response diversity a small negative effect
($r_\delta$= `r rpd`).

@@

Deriving total effects from the sum of direct and indirect effects in the
structural equation model (See Method), we showed that both environmental
stochasticity and species response diversity have by far the strongest total
effects on community stability (resp. $r_\delta$= `r rtsenv` and `r rtsdiv`,
Fig. \@ref(fig:sem)b). Species richness and average trophic level have similar
small but positive total effects on community stability ($r_\delta$=`r rtsr`),
but through opposite pathways. While species richness increases community
stability through a strong positive effect on asynchrony ($r_\delta$=`r rtar`),
it also weakly decreases population stability ($r_\delta$=`r rtpr`). In
contrast, average trophic level increases community stability through strong
positive effect on population stability ($r_\delta$=`r rtpt`), but is dampened
by a negative total effect on asynchrony ($r_\delta$=`r rtat`). Interestingly,
average interaction strength also has a positive effect on population stability
and a negative one on asynchrony (resp. $r_\delta$= `r rtpi` and `r rtai`). This
results in a total negative effect on community stability ($r_\delta$=
`r rtsi`). Connectance, instead, has a negative effect on community stability
($r_\delta$= `r rtsct`) through negative effects on both asynchrony and
population stability (resp. $r_\delta$=`r rtact` and `r rtpct`).

@@

)
pred_cap <- paste0("Slope of the community stability - species
  richness relationship based on food-web structure and response diversity. a,
  Diagrams displaying the slope coefficients of the effect of species richness
  on community stability. The average values of the control variables were used
  to generate the predictions: ", pv_cap, ". b, c: Examples of predictions
  from the model for two combinations of average trophic level and response
  diversity (values displayed in panel a). Lines display the mean predictions of
  the model. The coefficients of the linear model are displayed in Fig. S3.
  Relationships between population stability, asynchrony and species richness
  are displayed in Fig. S2.
  "
  ) %>%
  str_replace_all("\\s+", " ")
```

@@

We further assessed how environmental stochasticity and response diversity
interacted with food-web structural properties to determine the sign of the
stability-richness relationship, using a linear model including main effects and
all interactions among food-web structural properties, response diversity and
environmental stochasticity (see Methods). We found that there are strong
interactions between response diversity and food-web structure, which jointly
determine the sign of the stability-richness relationship.

In the absence of response diversity, we found only negative stability-richness
relationships, regardless of the food-web structure (Fig. \@ref(fig:pred)a).
Response diversity enables the rise of positive stability-richness
relationships, enhanced by average interaction strength and dampened by average
trophic level and connectance. Higher interaction strength leads to positive
stability-richness relationships for lower values of response diversity, and
interaction strength enhances asynchrony-richness relationships (Fig. S2), so
that the stronger positive stability-richness relationships are found at both
high interaction strength and high response diversity levels. In contrast,
higher average trophic level and connectance both lead to negative
stability-richness relationships, because they drive more negative population
stability-richness relationships and weaker positive asynchrony-richness
relationships (Fig. S2). Overall, our results highlight a strong contrast
between the relatively small effects of food-web structural properties on
overall community stability (Fig. \@ref(fig:sem)b) but their strong effects on
the slope of the stability-richness relationship (Fig. \@ref(fig:pred)a).

@@

With the projected changes in climate and land use, and the expected resulting
changes in community composition, it is important to understand the mechanisms
which allow ecological communities to remain stable despite perturbations. The
two components of temporal community stability are the average stability of the
populations composing the community on the one hand, and the asynchrony of the
population fluctuations on the other hand[@thibaut_understanding_2013;
@zhao_biodiversity_2022]. In simplified assemblages, such as single trophic
level communities, community stability has been shown to be driven by asynchrony
in species fluctuations, rather than by population
stability[@olivier_urbanization_2020]. However, whether this holds for more
complex ecological communities, such as food webs, remains largely unknown.
Here, we tackled this question using simulations of complex food-web dynamics
with a stochastic bioenergetic model. Our model results show that, in complex
food-webs, community stability is more driven by population stability than by
asynchrony, in contrast with what was observed in single trophic level
communities but in agreement with a previous food-web data
study[@danet_species_2021]. Our results thereby contribute to reconcile models
and data analyses. The study further reveals that the main drivers of community
stability are environmental stochasticity and response diversity, respectively
acting on population stability and asynchrony. Surprisingly, food-web structural
properties have in turn much weaker effects, a result consistent across
methodological choices (Fig. S1). Structural properties were however found to
play an important role in the sign of the stability-richness relationship, which
is strongly mediated by the interaction between food-web structure and response
diversity. These results highlight the importance of response diversity to
improve our understanding of the stability of ecological communities in variable
environments.

We found environmental stochasticity and response diversity to be the main
overall drivers of community stability. While environmental stochasticity
decreases population stability, response diversity generates asynchrony, thereby
buffering environmental stochasticity. This result is in accordance with
previous theoretical and empirical findings in competitive
communities[@thibaut_diversity_2012; @thibaut_understanding_2013;
@thebault_trophic_2005; @loreau_species_2008; @de_mazancourt_predicting_2013;
@ives_stability_2000]. In turn, response diversity was found to operate on
asynchrony mainly through portfolio effects in agreement with empirical findings
in competitive assemblages[@thibaut_diversity_2012; @zhao_biodiversity_2022].
This suggests that response diversity generates asynchrony mainly by the
statistical averaging of random independent species fluctuations[@doak_statistical_1998].

A consequence of the large effect of response diversity is that asynchrony in
species fluctuations is driven more by the statistical averaging of independent
species fluctuations (portfolio effects) than by compensatory dynamics emerging
from species interactions. This finding is also coherent with previous studies
in competitive communities[@thibaut_diversity_2012; @zhao_biodiversity_2022],
but it was less expected to happen in food-webs. Trophic interactions are indeed
expected to result in more compensatory dynamics resulting from oscillations
between predators and prey[@vandermeer_coupled_2004] or prey switching from
predators[@mccann_weak_1998; @mccann_diversitystability_2000]. However,
pioneering theoretical studies have shown that environmental stochasticity can
dampen prey switching[@vasseur_environmental_2007] and that trophic cascades
between trophic levels in species rich food-webs are weak compared to
food-chains and species poor food-web modules[@fahimipour_compensation_2017].
So, the question of the importance of compensatory dynamics in complex food webs
remains an open and interesting avenue for future research.

In contrast to empirical evidence in competitive
communities[@olivier_urbanization_2020], population stability (not asynchrony)
was found to be  the dominant driver of community stability in our food web
simulations. These results are in agreement with previous findings in empirical
food-webs[@danet_species_2021] and food-web
models[@eschenbrenner_diversity_2023], which suggests fundamental differences
between competitive communities and food webs in the relative contributions of
population stability and asynchrony to community stability. Our findings suggest
that the presence of trophic links contributes to synchronise the species of the
community because we found that the food-web metrics studied – namely average
trophic level, connectance, and average interaction strength – all decrease
asynchrony. These results are in line with empirical evidence that predators can
synchronise their prey[@raimondo_interspecific_2004;
@korpimaki_predatorinduced_2005], and that species pairs involved in trophic
interactions are more synchronous than those that are not[@danet_species_2021].
A complementary explanation for the relatively lower importance of asynchrony in
food webs (compared to single trophic level communities) could be related to the
biomass structure of food-webs. Food-webs typically include a larger range of
body masses than single trophic level assemblages[@brose_predator_2019]. Such
highly uneven biomass distribution can translate in uneven species temporal
variability, thereby creating uneven species contributions to
asynchrony[@thibaut_understanding_2013] and limiting portfolio effects and
compensatory dynamics in species-rich communities[@zhao_biodiversity_2022].

Investigating in more detail the drivers of community stability, our results
suggest that species richness and the network structural properties investigated
only have a weak effect on overall community stability. We found that higher
average trophic level leads to higher community stability by increasing average
population stability, confirming previous theoretical[@shanafelt_stability_2018;
@eschenbrenner_diversity_2023; @brose_allometric_2006] and
empirical[@danet_species_2021] results. However, this effect disappeared when we
simulated community dynamics with equal mortality rates for all species instead
of allometric ones (Fig. S1). This suggests that the higher stability of
food-webs with higher average trophic levels stems more from their lower
mortality rates rather than from their trophic role directly. The overall effect
of connectance on overall community stability was found to be quite low, in
agreement with previous empirical and theoretical findings[@danet_species_2021;
@eschenbrenner_diversity_2023]. Similarly, we found very low effects of the
average interaction strength on community stability, which is in apparent
disagreement with the predominant role of interaction strength and connectance
on stability[@may_will_1972].
